---
title: "JSON data wrangling"
author: "Victor Felix"
date: "2026-03-03"
output: html_document
---

```{r}
library(jsonlite)
library(stringr)
library(ggplot2)
library(dplyr)
library(stringr)
library(quanteda)
library(quanteda.textplots)
library(dplyr)
library(stringr)
library(tidyr)
```

Let's load the data and the #hashtags dictionary.

```{r}
# Read JSON file
data <- fromJSON("project-9858-comments_2026-01-21.json", flatten = TRUE)
dictionary <- read.csv("hashtag_dictionary.csv")
```

Here we subset the data so we only use images that had comments.

```{r}
# get only data with comments
commented_images<-data[data$board_title=="Notes" & data$comment_focus_type=="Subject",]
```

Here we subset images that had consensus based on the #expert_consensus tag.

```{r}
# Detect #expert_consensus on the comment body
expert_consensus<-str_detect(commented_images$comment_body, "#expert_consensus")

# Find images that had consensus based on "comment_focus_id" and expert_consensus == TRUE
images_with_consensus<-commented_images$comment_focus_id[expert_consensus]

# Add consensus (TRUE or FALSE) column to the commented images
commented_images$consensus <- 
  commented_images$comment_focus_id %in% images_with_consensus

head(commented_images)
```

Here we will get all tags from the comments body and add into a single tag column.

```{r}

# Function to get all %tags from each comment and include in a separate row
get_tags <- function(df) {
  df %>%
    mutate(
      tag = str_extract_all(comment_body, "#\\w+(?:_\\w+)*"),
      tag = sapply(tag, paste, collapse = ", ")
    )
}

commented_images_with_tags <- get_tags(commented_images)

head(commented_images_with_tags)

write.csv(commented_images_with_tags, file= "zoo_notes_comments.csv")
```

Sometimes people used more than one #tag in the same comment, so we will split these tags in 5 columns (tag1, tag2, tag3, tag4, tag5). 

```{r}
# Split tag column into multiple columns
# I had to add ,\\s* separator because there were spaces sometimes 

commented_images_with_tags <- commented_images_with_tags %>%
  separate(tag,
         into = paste0("tag", 1:5),
         sep = ",\\s*", # split at a comma followed by optional spaces
         fill = "right")

head(commented_images_with_tags)

```

Let's count the number of tags in each tag column  

```{r}
tag_cols <- c("tag1", "tag2", "tag3", "tag4", "tag5")

colSums(!is.na(commented_images_with_tags[, tag_cols]))
```


We want to subset only the #tags present in the bird dictionary. 

To do this I will create Booleans (True/False) for each column based on the dictionary data. Later, I will set to NA all #tags that were False in the Boolean (not in the bird dictionary). This is a good approach to maintain values which may be valid in one tag column but not in the others.

```{r}
# create a bird dictionary
bird_dictionary <-dictionary[dictionary$Class == "Ave",]

# get bird tags
bird_tags <- unique(bird_dictionary$tag)

# create a bird dictionary
is_bird_tag1 <- commented_images_with_tags$tag1 %in% bird_tags
is_bird_tag2 <- commented_images_with_tags$tag2 %in% bird_tags
is_bird_tag3 <- commented_images_with_tags$tag3 %in% bird_tags
is_bird_tag4 <- commented_images_with_tags$tag4 %in% bird_tags
is_bird_tag5 <- commented_images_with_tags$tag5 %in% bird_tags

# Remove from each column the names that are not in the dictionary
commented_images_with_tags$tag1[!is_bird_tag1] <- NA
commented_images_with_tags$tag2[!is_bird_tag2] <- NA
commented_images_with_tags$tag3[!is_bird_tag3] <- NA
commented_images_with_tags$tag4[!is_bird_tag4] <- NA
commented_images_with_tags$tag5[!is_bird_tag5] <- NA

# let's create a new dataframe for clarity
commented_images_with_bird_tags <- commented_images_with_tags
```

Let's count the number of bird tags in each tag column  

```{r}
tag_cols <- c("tag1", "tag2", "tag3", "tag4", "tag5")

colSums(!is.na(commented_images_with_bird_tags[, tag_cols]))
```

We need to identify the images based on the combined #tags of different comments for the same comment_focus_id. First we subset only the bird data that had consensus from the previous bird data.

```{r}
# Select only the comments from images that had consensus
birds_with_consensus<-commented_images_with_tags[commented_images_with_tags$consensus==TRUE,]
birds_with_consensus

```

Let's count the number of bird tags by column in the bird data with consensus. 

```{r}
tag_cols <- c("tag1", "tag2", "tag3", "tag4", "tag5")

colSums(!is.na(birds_with_consensus[, tag_cols]))

# almost all of the previous data
```


Now we need to use these information to identify birds in each image. First, let's add columns with scientific names for each #hashtag column.


```{r}
birds_comments_with_consensus <- birds_with_consensus %>%
  
  left_join(
    dictionary %>% select(tag, scientific_name),
    by = c("tag1" = "tag")
  ) %>%
  rename(scientific_name_tag1 = scientific_name) %>%
  
  left_join(
    dictionary %>% select(tag, scientific_name),
    by = c("tag2" = "tag")
  ) %>%
  rename(scientific_name_tag2 = scientific_name) %>%
  
  left_join(
    dictionary %>% select(tag, scientific_name),
    by = c("tag3" = "tag")
  ) %>%
  rename(scientific_name_tag3 = scientific_name)

head(birds_comments_with_consensus)
```

Maybe the same species are being identified more than once in the same comment, so we will keep only one unique species name in a given comment. Let's also remove "Ave" if there is a more specific bird identification.

```{r}

# this function will do what is described above
# remove duplicate species names in the same comment
# remove Ave if there is another bird identification present

remove_duplicates_per_comment <- function(df, sci_name_cols) {
  # loop on each row
  for (i in seq_len(nrow(df))) {
    # extract scientific names in a given row
    x <- as.character(unlist(df[i, sci_name_cols]))
    # remove NAs
    x <- x[!is.na(x)]
    # remove "Ave" if there are other scientific names
    if (length(x) > 1) {
      x <- x[x != "Ave"]
    }
    # keep unique values
    x_unique <- unique(x)
    # make unique values have the same length as sci_name_cols
    # it just add some NAs, it is important consistency
    length(x_unique) <- length(sci_name_cols)
    # replace the scientific names in that given row
    df[i, sci_name_cols] <- x_unique
  }
  return(df)
}

# let's use the names of the columns we want to evaluate
sci_name_cols_ <- paste0("scientific_name_tag", 1:3)

# apply the function
# note that the scientific names will not correspond to the tags anymore
# I think this if fine for or final analysis
birds_comments_with_consensus <- 
  remove_duplicates_per_comment(
    birds_comments_with_consensus,
    sci_name_cols = sci_name_cols_
  )

```

Let's count the number of unique scientific names across all comments. 

```{r}

colSums(!is.na(birds_comments_with_consensus[, sci_name_cols_]))

```

Here we create a new dataframe with comment_focus_id and the scientific names that we subset based on species mentioned in different comments. 

```{r}
unique_birds_per_comment <-birds_comments_with_consensus %>%
  select(comment_focus_id, starts_with("scientific")) %>%
  pivot_longer(
    cols = starts_with("scientific"),
    names_to = "source_column",
    values_to = "scientific_name"
  ) %>%
  filter(!is.na(scientific_name)) %>%
  select(comment_focus_id, scientific_name)
```

Here we identify the species if they had their names repeated at least twice.

```{r}
# count how many scientific names 
identified_birds_by_image <- unique_birds_per_comment %>%
  group_by(comment_focus_id, scientific_name) %>%
    summarise(n_id = n())

dual_expert_subjects <- identified_birds_by_image %>%
  filter(n_id > 1) 

#visualize
head(dual_expert_subjects)

# saving
write.csv(dual_expert_subjects,file="dual_expert_subjects.csv")
```
Plot the identified species

```{r}
# Barplot
ggplot(dual_expert_subjects, aes(x = scientific_name)) +
  geom_bar(fill = "steelblue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Repeated Hashtags per comment_focus_id", y = "Count", x = "Hashtag")

# number of species identified
NROW(unique(dual_expert_subjects$scientific_name))
```

A different plot a better visualization.

```{r}
# Create a vector of hashtags
scientific_name_vector <- dual_expert_subjects$scientific_name
scientific_name_corpus <- corpus(scientific_name_vector)
scientific_name_tokens <- tokens(scientific_name_corpus, remove_punct = TRUE)
scientific_name_dfm <- dfm(scientific_name_tokens)

# Set seed 
set.seed(100)

# Plot 
textplot_wordcloud(scientific_name_dfm,
                   max_words = 75, # max number of names
                   color = RColorBrewer::brewer.pal(8, "Dark2"))

```